\Large\bf FlexibleSUSY 2.0: Extensions to investigate the phenomenology of SUSY and non-SUSY models

### 2017-12-20 doc/flexiblesusy-2.0.tex

\begin{keyword}
sparticle,
supersymmetry,
Higgs,
renormalization group equations
\PACS 12.60.Jv
\PACS 14.80.Ly
\end{keyword}
\end{frontmatter}

@@

\subsection{Downloading \fs and generating a first spectrum generator}
\label{sec:quick-start-cmssm}
\fs is available as gzipped tarball on
[\url{http://flexiblesusy.hepforge.org}] or
[\url{https://github.com/FlexibleSUSY/FlexibleSUSY}] under the version
control system \texttt{git}.
%
To download and install \fstwo as a gzipped tarball run at the command
line:
%
\begin{lstlisting}[language=bash]
$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.0.tar.gz
$ tar -xf FlexibleSUSY-2.0.0.tar.gz
$ cd FlexibleSUSY-2.0.0
\end{lstlisting}%% $
%
\fstwo is distributed with a huge selection of predefined models
including: the CMSSM (called \modelname{CMSSM}), the semi-constrained and
fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}), the
$R$-symmetric MSSM (\modelname{MRSSM}) \cite{Kribs:2007ac}, the NUHM \ESSM
(\modelname{E6SSM}) \cite{Athron:2007en}, high-scale SUSY (\HSSUSY)
\cite{Bagnaschi:2015pwa}, split-SUSY (\modelname{SplitMSSM})
\cite{Bagnaschi:2015pwa} and the THDM-II (\modelname{THDMII}).  See the
contents of \code{model_files/} for all predefined model files.

@@

The calculated spectrum can be printed in an SLHA-compatible format
using the \code{FS<model>ToSLHA[]} function.
%
\begin{example}
  For the CMSSM, an example output of \code{FSCMSSMToSLHA[handle]}
  could look like:
\begin{lstlisting}
Block SPINFO
     1   FlexibleSUSY
     2   2.0.0
     5   CMSSM
     9   4.11.0
Block MASS
   1000021     1.14735362E+03   # Glu
        24     8.03935152E+01   # VWm
   1000024     3.85016460E+02   # Cha(1)
   1000037     6.43924799E+02   # Cha(2)
        25     1.14835832E+02   # hh(1)
        35     7.13118731E+02   # hh(2)
...
\end{lstlisting}
\end{example}
%
In \secref{sec:application_HSSUSY} several examples can be found that
illustrate how to perform parameter scans and uncertainty
estimates using the \mathematica interface of \fs's spectrum
generators.

@@

\begin{OptionTable}{\fs model file switches to enable/disable
    model-specific higher-order contributions in the SM,
    MSSM, NMSSM and split-MSSM.}{tab:fbsm_options}[lllX]
  %
  \multicolumn{4}{c}{\textit{SM}}\\
  \midrule
  \code{UseHiggs2LoopSM} & \code{False} & \code{True} or \code{False}
  & 2-loop contributions $O(\at^2 + \at \as)$ to $M_h$
  in
  the SM \\
  %
  \code{UseHiggs3LoopSM} & \code{False} & \code{True} or \code{False} &
  3-loop contributions $O(\at^3 + \at^2 \as \allowbreak + \at \as^2)$ to $M_h$ in
  the SM \\
  %
  \code{UseSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
  3-loop RGEs in the SM \\
  %
  \code{UseYukawa3LoopQCD} & \code{Automatic} & \multilinecell[t]{\code{True} or \code{False}\\ or \code{Automatic}} &
  2-loop and 3-loop QCD contributions $O(\as^2 + \as^3)$ to the \MSbar $y_t$ in the SM \\
  %
  \code{UseSMAlphaS3Loop} & \code{False} & \code{True} or \code{False}
  & 2-loop and 3-loop QCD threshold corrections
  $O(\as^2 + \as^3)$
  to $\as$ in the SM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{MSSM}}\\
  \midrule
  %
  \code{UseHiggs2LoopMSSM} & \code{False} & \code{True} or \code{False} &
  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
  \ab)^2 \allowbreak + \atau^2)$ to $M_h$, $M_H$ and $M_A$ in
  the MSSM \\
  %
  \code{UseHiggs3LoopMSSM} & \code{False} & \code{True} or
  \code{False} & 3-loop contributions
  $O(\at\as^2 \allowbreak + \ab\as^2)$ to $M_h$ in
  the MSSM (requires \Himalaya) \\
  %
  \code{UseMSSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
  3-loop RGEs in the MSSM \\
  %
  \code{UseMSSMYukawa2Loop} & \code{False} & \code{True} or \code{False} &
  2-loop \SQCD contribution $O(\as^2)$ to the \DRbar $y_b$ and $y_t$ in the MSSM \\
  %
  \code{UseMSSMAlphaS2Loop} & \code{False} & \code{True} or
  \code{False} & 2-loop \SQCD contribution
  $O(\as^2 + \as\at + \as\ab)$
  to $\as$ in the MSSM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{NMSSM}}\\
  \midrule
  %
  \code{UseHiggs2LoopNMSSM} & \code{False} & \code{True} or \code{False} &
  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
  \ab)^2 \allowbreak + \atau^2)$ to $M_{h_i}$ and $M_{A_i}$ in
  the NMSSM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{split-MSSM}}\\
  \midrule
  %
  \code{UseHiggs3LoopSplit} & \code{False} & \code{True} or \code{False} &
  3-loop contributions $O(\at\as^2)$ to $M_h$, $M_H$ and
  $M_A$ in the split-MSSM \\
  %
  \bottomrule
\end{OptionTable}

@@

In the SM, 2-loop and leading 3-loop contributions to the Higgs pole
mass are known in the \MSbar scheme
\cite{Degrassi:2012ry,Martin:2014cxa,Vega:2015fna}.  In \fs the 2-loop
contributions of $O(\at^2 + \at \as)$
\cite{Degrassi:2012ry,Martin:2014cxa} and the 3-loop
$O(\at^3 + \at^2 \as + \at \as^2)$ contributions \cite{Martin:2014cxa}
can be taken into account to calculate the SM-like Higgs pole mass for
non-SUSY models with only one Higgs.  In order to enable these loop
contributions, the following switches must be set in the model file:
%
\begin{lstlisting}
UseHiggs2LoopSM = True;
UseHiggs3LoopSM = True;
\end{lstlisting}
%
The enabled 2-loop SM contributions of
$O(\at^2 + \at \as)$ to the Higgs mass read
%
\begin{align}
  (\Delta m_h^2)_{2L} &= (\Delta m_h^2)_{2L}^{(\at^2)} + (\Delta m_h^2)_{2L}^{(\at\as)} \,, \\
  (\Delta m_h^2)_{2L}^{(\at^2)} &=
  \frac{2 t}{(4\pi)^4} \left[
    -3 y_t^4 \left(3 \;\barlog^2(t) - 7 \;\barlog(t) + 2 + \frac{\pi^2}{3}\right)
  \right] , \label{eq:2L_Mh_atat}\\
  (\Delta m_h^2)_{2L}^{(\at\as)} &=
      \frac{g_3^2 y_t^2}{(4\pi)^4} \;\bigg[
        \frac{37}{3} p^2 - \frac{122 p^4}{135 t} - 4 (5 p^2 - 8 t) \;\barlog(t) -
        12 (p^2 - 8 t) \;\barlog^2(t) \bigg] \label{eq:2L_Mh_atas} \\
  &\stackrel{\mathmakebox[\widthof{=}]{p^2=0}}{=} \
  \frac{2 t}{(4\pi)^4} \left[
    16 g_3^2 y_t^2 \left(3 \;\barlog^2(t) + \barlog(t)\right)
  \right] \,.
\end{align}
%
In Eqs.~\eqref{eq:2L_Mh_atat} ff.\ the abbreviations
$t \equiv m_t^2$ and $\barlog(t) \equiv \log(t/Q^2)$ have been used
for brevity, analogous to the notation of Ref.~\cite{Martin:2014cxa}.  Note
that the momentum dependence is included in the 2-loop expression of
Eq.~\eqref{eq:2L_Mh_atas} in order to generate implicit 3-loop
$O(\at^2\as)$ terms.
%
The explicit 3-loop effective potential contributions of
$O(\at^3 + \at^2 \as + \at \as^2)$ to the
Higgs mass included by \fs read
%
\begin{align}
\begin{split}
  (\Delta m_h^2)_{3L} &=
  \frac{m_t^2}{(4\pi)^6} \Bigg[
     g_3^4 y_t^2 \left(248.122 + 839.197 \;\barlog(t) + 160 \;\barlog^2(t) - 736 \;\barlog^3(t)\right) \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
     + g_3^2 y_t^4 \left(2764.365 + 1283.716 \;\barlog(t) - 360 \;\barlog^2(t) + 240 \;\barlog^3(t)\right) \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
     + y_t^6 \Big(-3199.017 + 36 \;\barlog(h) - 2653.511 \;\barlog(t) + 756 \;\barlog(h) \;\barlog(t)\\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}\qquad~~
     + \frac{27}{2} \;\barlog^2(t) + 324 \;\barlog(h) \;\barlog^2(t) - 225 \;\barlog^3(t)\Big)
  \Bigg]
\end{split}
\end{align}
%
with $h\equiv m_h^2$.  Note that the 3-loop Higgs mass calculation in
the \MSbar scheme in \fs is only complete at $O(\at\as^2)$.  The
3-loop contributions of $O(\at^3 + \at^2 \as)$ are currently incomplete
because they would require the \MSbar top Yukawa coupling to be determined
from the top pole mass at the 2-loop $O(\at^2 + \at\as)$.
However, these 2-loop contributions to the top Yukawa coupling are
currently not available in \fs. Furthermore, the proper inclusion of
corrections to the Higgs pole mass at $O(\at^3)$
would require the extension of Eq.~\eqref{eq:2L_Mh_atat} by momentum
dependent $O(\at^2)$ terms. Alternatively, the evaluation of the Higgs mass at
the renormalization scale $Q^2 = t$ implies that the neglected
3-loop contributions at $O(\at^3 t)$ vanish
\cite{Martin:2014cxa}.\footnote{Note that terms of $O(\at^3 h)$ are
  neglected here and in Ref.\ \cite{Martin:2014cxa}.}
Likewise, Eq.~\eqref{eq:2L_Mh_atat} neglects contributions of $O(\at^2h)$ which are
subdominant in comparison to the implemented $O(\at^2t)$ corrections.
In contrast, the QCD corrections to the pole mass at 2-loop level, Eq.~\eqref{eq:2L_Mh_atas}, involve
terms proportional to the quartic Higgs coupling $\lambda$ up to $O(\as\at h^2/t)$, which are not
neglected here.

@@

The pure SM QCD contribution to the \MSbar top Yukawa
coupling is known up to the 3-loop level of $O(\as^3)$
\cite{Chetyrkin:1999qi,Melnikov:2000qh}.  This 3-loop QCD expression
can be taken into account in \fs to extract the \MSbar top Yukawa
coupling in the SM from the top pole mass $M_t$ by setting
the following switch in the model file:
%
\begin{lstlisting}
UseYukawa3LoopQCD = True;
\end{lstlisting}
%
By default, \code{UseYukawa3LoopQCD} is set to \code{Automatic}, which
means that the 3-loop QCD contribution to the top Yukawa coupling is
taken into account only if the model is \MSbar renormalized.  To take
this 3-loop correction into account at runtime, the following flags
should be set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    7   3         # threshold corrections loop order
   24   123111221 # individual threshold correction loop orders
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {thresholdCorrectionsLoopOrder -> 3,
                   thresholdCorrections -> 123111221}
    ...];
\end{lstlisting}
%
The 3-loop QCD contribution is combined with the full 1-loop and QCD
2-loop contribution as
%
\begin{align}
\begin{split}
  m_t(Q) &= M_t + \re\Sigma_{t}^S(p^2=M_t^2,Q) \\
  &\phantom{={}} + M_t \Big[ \re\Sigma_{t}^L(p^2=M_t^2,Q) +
    \re\Sigma_{t}^R(p^2=M_t^2,Q) \\
  &\phantom{={} + M_t \Big[}
    + \Delta m_t^{(1),\text{qcd}}(Q) + \Delta m_t^{(2),\text{qcd}}(Q) + \Delta m_t^{(3),\text{qcd}}(Q) \Big]
  \,,
\end{split}
\end{align}
%
where $\Sigma_t^{S,L,R}(p^2,Q)$ denote the scalar, left- and
right-handed parts of the \MSbar-renormalized 1-loop top quark self
energy, $M_t$ is the top pole mass and
%
\begin{align}
  \Delta m_t^{(1),\text{qcd}}(Q) &=
  -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \;\barlog(t)\right], \\
  \begin{split}
    \Delta m_t^{(2),\text{qcd}}(Q) &= \left(\Delta
      m_t^{(1),\text{qcd}}(Q)\right)^2 - \frac{g_3^4}{4608 \pi^4}
    \Bigg[396 \;\barlog^2(t) - 1452 \;\barlog(t) \\
    &\phantom{={}}
    - 48 \zeta(3) + 2053 + 16 \pi^2 (1+\log 4)\Bigg] ,
  \end{split} \\
\begin{split}
  \Delta m_t^{(3),\text{qcd}}(Q) &= \frac{g_3^6}{2430\, (4\pi)^6}
  \Bigg\{
  48600 \;\barlog^3(t) - 208980 \;\barlog^2(t) \\
  &\phantom{=\;} +540 \Big[-1560\,\zeta (3)+2993+40 \pi ^2 (1+\log 4)\Big] \barlog(t)\\
  &\phantom{=\;} +15 \left[69120\,\text{Li}_4\left(\frac{1}{2}\right)+113040\,\zeta(3)-
    94800 \,\zeta(5)-280853+2880 \,\log^4 2\right]\\
  &\phantom{=\;} +4 \pi^2 \Big[129510 \,\zeta(3)-388781 + 240 (733+24 \log 2) \,\log 2\Big] - 10500 \pi^4
  \Bigg\}
\end{split}\\
  &\approx
  \frac{g_3^6}{(4\pi)^6}\, 20 \left[\barlog^3(t)
    - \frac{43}{10} \;\barlog^2(t)
    + 22.8874 \;\barlog(t)
    -172.937 \right] \,.
\end{align}
%
In \figref{fig:SM_thresholds} the impact of the 3-loop correction to
the \MSbar top Yukawa coupling in the SM $y_t^{\SM}(M_Z)$
on the prediction of the light \CP-even Higgs pole mass in the MSSM in
the pure EFT calculation of \HSSUSY is shown as the green
dashed-double-dotted line.  As already discussed in
Ref.~\cite{Vega:2015fna}, we find that the inclusion of the 3-loop
correction to $y_t^{\SM}(M_Z)$ reduces the Higgs mass by up to
$500\unit{MeV}$.  Note that this is formally a (partial) 4-loop
effect on the Higgs mass, which is beyond the current accuracy of
\HSSUSY.

@@

In the MSSM, the 3-loop $\beta$ functions are also known
\cite{Jack:2003sx,Jack:2004ch}.  \fs allows the user to take these 3-loop
$\beta$ functions into account in the running of the \DRbar MSSM parameters by
setting the following switch in the model file:
%
\begin{lstlisting}
UseMSSM3LoopRGEs = True;
\end{lstlisting}
%
To use 3-loop running at runtime, the following flag should also be set in
the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    6   3         # beta-functions loop order
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration
corresponds to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {betaFunctionLoopOrder -> 3}
    ...];
\end{lstlisting}
%
The expressions for the 3-loop $\beta$ functions have been extracted
from the official FORM file provided by the authors of
Refs.~\cite{Jack:2003sx,Jack:2004ch}.\footnote{\url{https://www.liverpool.ac.uk/~dij/betas/allgennb.log}}
We have numerically compared the expressions with the ones implemented
in \softsusy 3.7.0 \cite{Allanach:2014nba} and found exact agreement.
%
The effect of the 3-loop RGEs on the Higgs pole mass in the MSSM is of
the order of a few $100\unit{MeV}$ as discussed in
Ref.~\cite{Allanach:2014nba} and is shown in \figref{fig:MSSMMuBMu}.
%
Due to the complexity of the 3-loop $\beta$ functions, the runtime of
the MSSM spectrum generators is increased by a factor of $4$--$5$ if
3-loop running is enabled in the MSSM\@.

@@

Print["Mh = (", Mh, " +- ", DMh, ") GeV"];
  \end{lstlisting}
  The output of the script could read
  \begin{lstlisting}
Mh = (125.047 \pm 1.52741) GeV
  \end{lstlisting}
\end{example}

@@

\tabref{tab:sm_mma_options}.  Unset parameters are set to their

>>>>

\begin{keyword}
sparticle,
supersymmetry,
Higgs,
renormalization group equations
\PACS 12.60.Cn, 12.60.Fr, 12.60.Jv, 12.60.-i
\PACS 14.80.Bn, 14.80.Da, 14.80.Ec, 14.80.Fd, 14.80.Ly, 14.80.Nb, 14.80.Sv
\end{keyword}
\end{frontmatter}

@@

\subsection{Downloading \fs and generating a first spectrum generator}
\label{sec:quick-start-cmssm}
\fs is available as gzipped tarball on
[\url{http://flexiblesusy.hepforge.org}] or
[\url{https://github.com/FlexibleSUSY/FlexibleSUSY}] under the version
control system \texttt{git}.
%
To download and install \fstwo as a gzipped tarball run at the command
line:
%
\begin{lstlisting}[language=bash]
$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.1.tar.gz
$ tar -xf FlexibleSUSY-2.0.1.tar.gz
$ cd FlexibleSUSY-2.0.1
\end{lstlisting}%% $
%
\fstwo is distributed with a huge selection of predefined models
including: the CMSSM (called \modelname{CMSSM}), the semi-constrained and
fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}), the
$R$-symmetric MSSM (\modelname{MRSSM}) \cite{Kribs:2007ac}, the NUHM \ESSM
(\modelname{E6SSM}) \cite{Athron:2007en}, high-scale SUSY (\HSSUSY)
\cite{Bagnaschi:2015pwa}, split-SUSY (\modelname{SplitMSSM})
\cite{Bagnaschi:2015pwa} and the THDM-II (\modelname{THDMII}).  See the
contents of \code{model_files/} for all predefined model files.

@@

The calculated spectrum can be printed in an SLHA-compatible format
using the \code{FS<model>ToSLHA[]} function.
%
\begin{example}
  For the CMSSM, an example output of \code{FSCMSSMToSLHA[handle]}
  could look like:
\begin{lstlisting}
Block SPINFO
     1   FlexibleSUSY
     2   2.0.1
     5   CMSSM
     9   4.11.0
Block MASS
   1000021     1.14735362E+03   # Glu
        24     8.03935152E+01   # VWm
   1000024     3.85016460E+02   # Cha(1)
   1000037     6.43924799E+02   # Cha(2)
        25     1.14835832E+02   # hh(1)
        35     7.13118731E+02   # hh(2)
...
\end{lstlisting}
\end{example}
%
In \secref{sec:application_HSSUSY} several examples can be found that
illustrate how to perform parameter scans and uncertainty
estimates using the \mathematica interface of \fs's spectrum
generators.

@@

\begin{OptionTable}{\fs model file switches to enable/disable
    model-specific higher-order contributions in the SM,
    MSSM, NMSSM and split-MSSM.}{tab:fbsm_options}[lllX]
  %
  \multicolumn{4}{c}{\textit{SM}}\\
  \midrule
  \code{UseHiggs2LoopSM} & \code{False} & \code{True} or \code{False}
  & 2-loop contributions $O(\at^2 + \at \as)$ to $M_h$ in the SM \\
  %
  \code{UseHiggs3LoopSM} & \code{False} & \code{True} or \code{False} &
  3-loop contributions $O(\at^3 + \at^2 \as \allowbreak + \at \as^2)$ to $M_h$ in
  the SM \\
  %
  \code{UseSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
  3-loop RGEs in the SM \\
  %
  \code{UseYukawa3LoopQCD} & \code{Automatic} & \multilinecell[t]{\code{True} or \code{False}\\ or \code{Automatic}} &
  2-loop and 3-loop QCD contributions $O(\as^2 + \as^3)$ to the \MSbar $y_t$ in the SM \\
  %
  \code{UseSMAlphaS3Loop} & \code{False} & \code{True} or \code{False}
  & 2-loop and 3-loop QCD threshold corrections
  $O(\as^2 + \as^3)$
  to the \MSbar $\as$ in the SM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{MSSM}}\\
  \midrule
  %
  \code{UseHiggs2LoopMSSM} & \code{False} & \code{True} or \code{False} &
  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
  \ab)^2 \allowbreak + \atau^2)$ to $M_h$, $M_H$ and $M_A$ in
  the MSSM \\
  %
  \code{UseHiggs3LoopMSSM} & \code{False} & \code{True} or
  \code{False} & 3-loop contributions
  $O(\at\as^2 \allowbreak + \ab\as^2)$ to $M_h$ in
  the MSSM (requires \Himalaya) \\
  %
  \code{UseMSSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
  3-loop RGEs in the MSSM \\
  %
  \code{UseMSSMYukawa2Loop} & \code{False} & \code{True} or \code{False} &
  2-loop \SQCD contribution $O(\as^2)$ to the \DRbar $y_b$ and $y_t$ in the MSSM \\
  %
  \code{UseMSSMAlphaS2Loop} & \code{False} & \code{True} or
  \code{False} & 2-loop \SQCD contribution
  $O(\as^2 + \as\at + \as\ab)$
  to the \DRbar $\as$ in the MSSM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{NMSSM}}\\
  \midrule
  %
  \code{UseHiggs2LoopNMSSM} & \code{False} & \code{True} or \code{False} &
  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
  \ab)^2 \allowbreak + \atau^2)$ to $M_{h_i}$ and $M_{A_i}$ in
  the NMSSM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{split-MSSM}}\\
  \midrule
  %
  \code{UseHiggs3LoopSplit} & \code{False} & \code{True} or \code{False} &
  3-loop contributions $O(\at\as^2)$ to $M_h$, $M_H$ and
  $M_A$ in the split-MSSM \\
  %
  \bottomrule
\end{OptionTable}

@@

In the SM, 2-loop and leading 3-loop contributions to the Higgs pole
mass are known in the \MSbar scheme
\cite{Degrassi:2012ry,Martin:2014cxa,Vega:2015fna}.  In \fs the 2-loop
contributions of $O(\at^2 + \at \as)$
\cite{Degrassi:2012ry,Martin:2014cxa} and the 3-loop
$O(\at^3 + \at^2 \as + \at \as^2)$ contributions \cite{Martin:2014cxa}
can be taken into account to calculate the SM-like Higgs pole mass for
non-SUSY models with only one Higgs.  In order to enable these loop
contributions, the following switches must be set in the model file:
%
\begin{lstlisting}
UseHiggs2LoopSM = True;
UseHiggs3LoopSM = True;
\end{lstlisting}
%
The enabled 2-loop SM contributions of
$O(\at^2 + \at \as)$ to the Higgs mass read
%
\begin{align}
  (\Delta m_h^2)_{2L} &= (\Delta m_h^2)_{2L}^{(\at^2)} + (\Delta m_h^2)_{2L}^{(\at\as)} \,, \\
  (\Delta m_h^2)_{2L}^{(\at^2)} &=
  \frac{2 t}{(4\pi)^4} \left[
    -3 y_t^4 \left(3 \;\barlog^2(t) - 7 \;\barlog(t) + 2 + \frac{\pi^2}{3}\right)
  \right] , \label{eq:2L_Mh_atat}\\
  (\Delta m_h^2)_{2L}^{(\at\as)} &=
      \frac{g_3^2 y_t^2}{(4\pi)^4} \;\bigg[
        \frac{37}{3} p^2 - \frac{122 p^4}{135 t} - 4 (5 p^2 - 8 t) \;\barlog(t) -
        12 (p^2 - 8 t) \;\barlog^2(t) \bigg] \label{eq:2L_Mh_atas} \\
  &\stackrel{\mathmakebox[\widthof{=}]{p^2=0}}{=} \
  \frac{2 t}{(4\pi)^4} \left[
    16 g_3^2 y_t^2 \left(3 \;\barlog^2(t) + \barlog(t)\right)
  \right] \,.
\end{align}
%
In Eqs.~\eqref{eq:2L_Mh_atat} ff.\ the abbreviations
$t \equiv m_t^2$ and $\barlog(t) \equiv \log(t/Q^2)$ have been used
for brevity, analogous to the notation of Ref.~\cite{Martin:2014cxa},
where $m_t$ is the \MSbar top mass in the SM\@.  Note
that the momentum dependence is included in the 2-loop expression of
Eq.~\eqref{eq:2L_Mh_atas} in order to generate implicit 3-loop
$O(\at^2\as)$ terms.
%
The explicit 3-loop effective potential contributions of
$O(\at^3 + \at^2 \as + \at \as^2)$ to the
Higgs mass included by \fs read
%
\begin{align}
  (\Delta m_h^2)_{3L} &=
  \frac{m_t^2}{(4\pi)^6} \Bigg[
     g_3^4 y_t^2 \left(248.122 + 839.197 \;\barlog(t) + 160 \;\barlog^2(t) - 736 \;\barlog^3(t)\right) \nonumber \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
     + g_3^2 y_t^4 \left(2764.365 + 1283.716 \;\barlog(t) - 360 \;\barlog^2(t) + 240 \;\barlog^3(t)\right) \nonumber \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
     + y_t^6 \Big(-3199.017 + 36 \;\barlog(h) - 2653.511 \;\barlog(t) + 756 \;\barlog(h) \;\barlog(t) \nonumber \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}\qquad~~
     + \frac{27}{2} \;\barlog^2(t) + 324 \;\barlog(h) \;\barlog^2(t) - 225 \;\barlog^3(t)\Big)
  \Bigg]
\end{align}
%
where $h\equiv m_h^2$ is the squared \MSbar Higgs mass in the SM\@.
Note that the 3-loop Higgs mass calculation in
the \MSbar scheme in \fsbreak is only complete at $O(\at\as^2)$.  The
3-loop contributions of $O(\at^3 + \at^2 \as)$ are currently incomplete
because they would require the \MSbar top Yukawa coupling to be determined
from the top pole mass at the 2-loop $O(\at^2 + \at\as)$.
However, these 2-loop contributions to the top Yukawa coupling are
currently not available in \fs. Furthermore, the proper inclusion of
corrections to the Higgs pole mass at $O(\at^3)$
would require the extension of Eq.~\eqref{eq:2L_Mh_atat} by momentum
dependent $O(\at^2)$ terms. Alternatively, the evaluation of the Higgs mass at
the renormalization scale $Q^2 = t$ implies that the neglected
3-loop contributions at $O(\at^3 t)$ vanish
\cite{Martin:2014cxa}.\footnote{Note that terms of $O(\at^3 h)$ are
  neglected here and in Ref.\ \cite{Martin:2014cxa}.}
Likewise, Eq.~\eqref{eq:2L_Mh_atat} neglects contributions of $O(\at^2h)$ which are
subdominant in comparison to the implemented $O(\at^2t)$ corrections.
In contrast, the QCD corrections to the pole mass at 2-loop level, Eq.~\eqref{eq:2L_Mh_atas}, involve
terms proportional to the quartic Higgs coupling $\lambda$ up to $O(\as\at h^2/t)$, which are not
neglected here.

@@

The pure SM QCD contribution to the \MSbar top Yukawa
coupling is known up to the 3-loop level of $O(\as^3)$
\cite{Chetyrkin:1999qi,Melnikov:2000qh}.  This 3-loop QCD expression
can be taken into account in \fs to extract the \MSbar top Yukawa
coupling in the SM from the top pole mass $M_t$ by setting
the following switch in the model file:
%
\begin{lstlisting}
UseYukawa3LoopQCD = True;
\end{lstlisting}
%
By default, \code{UseYukawa3LoopQCD} is set to \code{Automatic}, which
means that the 3-loop QCD contribution to the top Yukawa coupling is
taken into account only if the model is \MSbar renormalized.  To take
this 3-loop correction into account at runtime, the following flags
should be set in the SLHA input file:
%
\\\begin{minipage}{\linewidth}
\begin{lstlisting}
Block FlexibleSUSY
    7   3         # threshold corrections loop order
   24   123111221 # individual threshold correction loop orders
\end{lstlisting}
\end{minipage}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {thresholdCorrectionsLoopOrder -> 3,
                   thresholdCorrections -> 123111221}
    ...];
\end{lstlisting}
%
The 3-loop QCD contribution is combined with the full 1-loop and QCD
2-loop contribution as
%
\begin{align}
\begin{split}
  m_t(Q) &= M_t + \re\Sigma_{t}^S(p^2=M_t^2,Q) \\
  &\phantom{={}} + M_t \Big[ \re\Sigma_{t}^L(p^2=M_t^2,Q) +
    \re\Sigma_{t}^R(p^2=M_t^2,Q) \\
  &\phantom{={} + M_t \Big[}
    + \Delta m_t^{(1),\text{qcd}}(Q) + \Delta m_t^{(2),\text{qcd}}(Q) + \Delta m_t^{(3),\text{qcd}}(Q) \Big]
  \,,
\end{split} \label{eq:mt_MSbar}
\end{align}
%
where $\Sigma_t^{S,L,R}(p^2,Q)$ denote the scalar, left- and
right-handed parts of the \MSbar-renormalized 1-loop top quark self
energy, $M_t$ is the top pole mass and
%
\begin{align}
  \Delta m_t^{(1),\text{qcd}}(Q) &=
  -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \;\barlog(t)\right], \\
  \begin{split}
    \Delta m_t^{(2),\text{qcd}}(Q) &= \left(\Delta
      m_t^{(1),\text{qcd}}(Q)\right)^2 - \frac{g_3^4}{4608 \pi^4}
    \Bigg[396 \;\barlog^2(t) - 1452 \;\barlog(t) \\
    &\phantom{={}}
    - 48 \zeta(3) + 2053 + 16 \pi^2 (1+\log 4)\Bigg] ,
  \end{split} \\
\begin{split}
  \Delta m_t^{(3),\text{qcd}}(Q) &= \frac{g_3^6}{2430\, (4\pi)^6}
  \Bigg\{
  48600 \;\barlog^3(t) - 208980 \;\barlog^2(t) \\
  &\phantom{=\;} +540 \Big[-1560\,\zeta (3)+2993+40 \pi ^2 (1+\log 4)\Big] \barlog(t)\\
  &\phantom{=\;} +15 \left[69120\,\text{Li}_4\left(\frac{1}{2}\right)+113040\,\zeta(3)-
    94800 \,\zeta(5)-280853+2880 \,\log^4 2\right]\\
  &\phantom{=\;} +4 \pi^2 \Big[129510 \,\zeta(3)-388781 + 240 (733+24 \log 2) \,\log 2\Big] - 10500 \pi^4
  \Bigg\}
\end{split}\\
  &\approx
  \frac{g_3^6}{(4\pi)^6}\, 20 \left[\barlog^3(t)
    - \frac{43}{10} \;\barlog^2(t)
    + 22.8874 \;\barlog(t)
    -172.937 \right] \,.
\end{align}
%
In \figref{fig:SM_thresholds} the impact of the 3-loop correction to
the \MSbar top Yukawa coupling in the SM $y_t^{\SM}(M_Z)$
on the prediction of the light \CP-even Higgs pole mass in the MSSM in
the pure EFT calculation of \HSSUSY is shown as the green
dashed-double-dotted line.  As already discussed in
Ref.~\cite{Vega:2015fna}, we find that the inclusion of the 3-loop
correction to $y_t^{\SM}(M_Z)$ reduces the Higgs mass by up to
$500\unit{MeV}$.  Note that this is formally a (partial) 4-loop
effect on the Higgs mass, which is beyond the current accuracy of
\HSSUSY.

@@

In the MSSM, the 3-loop $\beta$ functions are also known
\cite{Jack:2003sx,Jack:2004ch}.  \fs allows the user to take these 3-loop
$\beta$ functions into account in the running of the \DRbar MSSM parameters by
setting the following switch in the model file:
%
\begin{lstlisting}
UseMSSM3LoopRGEs = True;
\end{lstlisting}
%
To use 3-loop running at runtime, the following flag should also be set in
the SLHA input file:
%
\\\begin{minipage}{\linewidth}
\begin{lstlisting}
Block FlexibleSUSY
    6   3         # beta-functions loop order
\end{lstlisting}
\end{minipage}
%
In \fs's \mathematica interface, the above SLHA configuration
corresponds to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {betaFunctionLoopOrder -> 3}
    ...];
\end{lstlisting}
%
The expressions for the 3-loop $\beta$ functions have been extracted
from the official FORM file provided by the authors of
Refs.~\cite{Jack:2003sx,Jack:2004ch}.\footnote{\url{https://www.liverpool.ac.uk/~dij/betas/allgennb.log}}
We have numerically compared the expressions with the ones implemented
in \softsusy 3.7.0 \cite{Allanach:2014nba} and found exact agreement.
%
The effect of the 3-loop RGEs on the Higgs pole mass in the MSSM is of
the order of a few $100\unit{MeV}$ as discussed in
Ref.~\cite{Allanach:2014nba} and is shown in \figref{fig:MSSMMuBMu}.
%
Due to the complexity of the 3-loop $\beta$ functions, the runtime of
the MSSM spectrum generators is increased by a factor of $4$--$5$ if
3-loop running is enabled in the MSSM\@.

@@

Print["Mh = (", Mh, " +- ", DMh, ") GeV"];
  \end{lstlisting}
  The output of the script could read
  \begin{lstlisting}
Mh = (125.047 +- 1.52741) GeV
  \end{lstlisting}
\end{example}

@@

\tabref{tab:sm_mma_options}.\footnote{Note that in \fs's \mathematica
  interface, the fine structure constant $\aem^{\SM(5)}(M_Z)$ is input,
  not its inverse as in the SLHA standard.}
Unset parameters are set to their
